---
title: "Global_Space_Exploration_Analysis"
author: "Ifta kakazai"
date: "2025-05-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 📘 Project Summary: Global Space Exploration Analysis in R

This project presents a comprehensive exploratory data analysis of a dataset on global space exploration activities using R programming. The dataset includes key information such as; 
  1:  Launching country, 
  2:  Mission year, 
  3:  Mission type, 
  4:  Launch site, 
  5:  Satellite type, 
  6:  Budget, 
  7:  Success rate, 
  8:  Technology used, 
  9:  Environmental impact, 
  10: Collaborating countries &,
  11: Mission duration. 

The primary goal was to derive meaningful insights from this data while practicing and demonstrating proficiency in data wrangling and data visualization using the tidyverse, particularly dplyr and ggplot2.


## Packages Including

```{r}
library(dplyr)
library(ggplot2)
library(readr)
```

## Importing the CSV file

```{r}
Dataset <- read_csv("C:/Users/ifta/OneDrive/Desktop/R Directory/Data Analysis Project/Global_Space_Exploration_Dataset.csv")

```

## Detailed Summary of the Dataset

```{r}
summary(Dataset)
```


# Data Analysis


## Country which has launched the most space missions overall.

```{r}
top_mission_countries <- Dataset %>%
  count(Country, name = "Total_Missions") %>%
  arrange(desc(Total_Missions)) %>%
  slice_max(Total_Missions, n=3)
  
print(top_mission_countries)   
```

## Global success rate of missions changed over the years.
```{r}
Global_Success_Rate <- Dataset %>%
                          filter(Year %in% c(2021,2022,2023,2024,2025)) %>%
                          group_by(Year) %>%                          
                          summarize(avg_success_rate = mean(Success_Rate))
ggplot(Global_Success_Rate, aes(x=Year, y= avg_success_rate)) +
  geom_line(color = "grey") +
  geom_point(color = "black") +
  labs(title = "Global Average Success Rate of Missions Over the Years", x = "Years", y = "Average Success Rate (%)") +
  theme_minimal()
  
```

## Top 5 countries that have spent the most on space missions.
```{r}
Top_Countries_spending <- Dataset %>%
group_by(Country) %>%
summarize(Total_Budget_in_Billion = sum(`Budget_(in Billion $)`)) %>%
slice_max(Total_Budget_in_Billion, n=5)

print(Top_Countries_spending)
  
```

## Satellite types that were most commonly used across all missions.

```{r}
Common_Satellites_Type <- Dataset %>%
  group_by(Satellite_Type) %>%
  count(Satellite_Type, name = 'Number_of_Missions', sort = TRUE)

print(Common_Satellites_Type)

```

## The distribution of mission durations.

```{r}

ggplot(na.omit(Dataset), aes(x = `Duration_(in Days)`)) +
  geom_histogram(bins = 30, fill = "#748cab", color = "white") +
  labs(title = "Distribution of Mission Durations",
       x = "Duration (in Days)",
       y = "Number of Missions") +
  theme_minimal()
  
```




## Comparison of the budgets and missions by country.

```{r}
Comparison_budget_country <- Dataset %>%
  select(Country, `Budget_(in Billion $)`) %>%
  filter(Country %in% c("China","UK", "France", "Israel"))


ggplot(Comparison_budget_country, aes(x= Country, y= `Budget_(in Billion $)`)) +
  geom_boxplot(color = c("China" = "#748cab",
                         "UK" = "#3e5c76",
                         "France" = "#fca311",
                         "Israel" = "#e5e5e5")) +
  labs(title = "Budgets of mission by Country", 
         x= "Country",
         y= "Budgeted Amount in Billion") +
  theme_minimal()
```



## Relationship between budget and success rate.

```{r}
ggplot(Dataset, aes(x=`Budget_(in Billion $)`
                    , y= Success_Rate)) +
   geom_jitter(color = "steelblue", width = 0.1, height = 1, alpha = 0.6) +
  scale_x_log10() +
  labs(title = "Budget vs. Mission Success Rate",
       x = "Budget in Billion $",
       y = "Success Rate (%)") +
  theme_minimal()
```






## The number of missions for each Mission Type.

```{r}
Number_of_missions_and_type <- Dataset %>%
  count(`Mission Type` , name = "Number_of_Missions") %>%
  group_by(`Mission Type`)

ggplot(Number_of_missions_and_type, aes(x= `Mission Type`, y= Number_of_Missions)) +
  geom_col(alpha = 0.01, color = c("Manned" = "#14213d",
                                             "Unmanned" = "#3a7ca5")) + 
  labs(title = "Distribution of Missions by Type: Manned vs Unmanned")+
  theme_minimal()
```



## Number of Missions by Sattellite Type.

```{r}
Missions_and_Satellite_type <- Dataset %>%
  count(Satellite_Type , name = "Number_of_Missions") %>%
  group_by(Satellite_Type)



ggplot(Missions_and_Satellite_type, aes(x= Satellite_Type, y= Number_of_Missions)) +
  geom_col(fill = c("Communication" = "#0d1321",
                               "Navigation" = "#1d2d44",
                               "Research" = "#3e5c76",
                               "Spy" = "#748cab",
                               "Weather" = "#f0ebd8")) +
  labs(title = "Number of Mission Vs. Satellite Type", x= "Satellite Type", y = "Number of Missions") +
  theme_minimal()
```



## Top Chinese Space Missions by Budget in 2025

```{r}
Top_Budgeted_Mission <- Dataset %>%
    filter(Country == "China", 
           Year == 2025,
           `Budget_(in Billion $)`>=20.0) %>%
    select(Mission_Name,`Budget_(in Billion $)`,Success_Rate)%>%
    arrange(desc(`Budget_(in Billion $)`))
    

print(Top_Budgeted_Mission)
```








